QPLIB: A Library of Quadratic Programming Instances

### 2017-01-20 QPLIB-2.3.tex
start revising solution methods

The methods implemented by the solvers in Table~\ref{t:solvers} can, following \cite{neumaier}, be broadly organized in three categories: {\it incomplete}, {\it asymptotically complete}, and {\it complete}. Incomplete methods are only able to identify solutions, often locally optimal according to a suitable notion, and may even fail to find one even when one exists; in particular, they are typically not able to determine that an instance is empty. Asymptotically complete methods can find a globally optimal solution with probability one in infinite time, but again they cannot prove that a given instance is infeasible. Complete methods find an approximate globally optimal solution to within a prescribed optimality tolerance within finite time, or prove that none such exists (but see Sect.~\ref{s:complete} below); they are often referred to as \emph{exact} methods in the computational optimization community. According to \cite{neumaier}, {\it rigorous} methods also exist which find global within given tolerances even in the presence of rounding errors, except in ``near-degenerate cases''. Since it is debatable whether any of the solver we are using can be classified as rigorous, we have elected to limit ourselves to declaring solvers complete.

Incomplete methods are usually realized as local search algorithms, asymptotically complete methods are usually realized by meta-heuristic methods such as multi-start or simulated annealing, and complete methods for $\mathcal{NP}$-hard problems such as QP are typically realized as implicit exhaustive exploration algorithms. However, these three categories may exhibit some overlap. For example, any deterministic method for solving QCQ locally is incomplete in general, but becomes complete for CCC, since any local optimum of a convex QP is also global. Therefore, when we state that a given algorithm is incomplete or (asymptotically) complete we mean that it is so the largest problem class that the solver naturally targets, although it may be complete on specific sub-classes. For example, {\sc SnOpt} naturally targets continuous nonconvex NLPs and it is incomplete on this class, and therefore on QCQ, but becomes complete for CCC. In general, all complete methods for a problem $P$ must be complete for any problem $Q \subseteq P$, while a complete method for $P$ might be incomplete for $R \supset P$.

@@

Local search methods typically require as an input a solution $x'$ and attempt to improve it---either towards feasibility, or towards optimality, or both---using only information that is available in a neighborhood of $x'$. In general, local search methods are incomplete. As remarked above, however, local search methods deployed on a convex problem behave like complete methods, possibly via devices that that allow one to find a feasible starting solution if there is one (like what was used to be called ``phase -1'' in the simplex method).

Most local search methods for *C* are iterative in nature, and need a feasible starting point $x^0$ as input. The ($k+1$)-st iterate is obtained as $x^{k+1} = x^k + \alpha_k d^k$, where $\alpha_k$ (a scalar) is the {\it step length}, and $d^k$ (a vector) is the {\it search direction}, e.g.~belonging to a tangent manifold and having a negative directional derivative at $x^k$, in order to improving optimality while reducing infeasibility. Alternatively, $x^{k+1} = x^k + d^k(\alpha_k)$ where $d^k(\cdot)$ is an arc satisfying $d^k(0) = 0$. Local search methods for *I* are also iterative, but since defining a useful notion of descent direction is harder in presence of integer variables, $d^k$ and $\alpha_k$ are usually computed in different ways, depending on the application at hand.

The solvers in Table \ref{t:solvers} which implement incomplete methods are {\sc ConOpt}, {\sc IpOpt}, {\sc Minos}, {\sc SnOpt} and {\sc KNitro}. The former four solvers implement local search algorithms for continuous nonconvex NLPs (a problem class containing QCQ). Note that traditionally {\sc KNitro} used to be a local (nonconvex) NLP solver. All of these solvers implement essentially three types of local search methods for NLP:
%
\begin{itemize}
 \item active set methods ({\sc ConOpt}, {\sc Minos}, {\sc SnOpt}) \cite{Dost09};
 %
 \item interior point methods ({\sc IpOpt}, {\sc KNitro}) \cite{Wright97};
 %
 \item projected gradient methods \cite{CalaMore87:mp,ChenGui13:coap}.
\end{itemize}
%
Active set and interior point methods have been defined for *C* but also for general (continuous) NLPs. In fact, all of the solvers named above target this larger problem class.

>>>>

The methods implemented by the solvers in Table~\ref{t:solvers} can, following \cite{neumaier}, be broadly organized in three categories: {\it incomplete}, {\it asymptotically complete}, and {\it complete}. Incomplete methods are only able to identify solutions, often locally optimal according to a suitable notion, and may even fail to find one even when one exists; in particular, they are typically not able to determine that an instance is empty. Asymptotically complete methods can find a globally optimal solution with probability one in infinite time, but again they cannot prove that a given instance is infeasible. Complete methods find an approximate globally optimal solution within a prescribed optimality tolerance within finite time, or prove that none such exists (but see Sect.~\ref{s:complete} below); they are often referred to as \emph{exact} methods in the computational optimization community. According to \cite{neumaier}, {\it rigorous} methods also exist, which find globally optimal solutions within given tolerances even in the presence of rounding errors, except for ``near-degenerate cases''. Since none of the solvers we are using can be classified as rigorous, we limit ourselves to declaring solvers complete.

Incomplete methods are usually realized as local search algorithms, asymptotically complete methods are usually realized by meta-heuristic methods such as multi-start or simulated annealing, and complete methods for $\mathcal{NP}$-hard problems such as QP are typically realized as implicit exhaustive exploration algorithms. However, these three categories may exhibit some overlap. For example, any deterministic method for solving QCQ locally is incomplete in general, but becomes complete for CCC, since any local optimum of a convex QP is also global. Therefore, when we state that a given algorithm is incomplete or (asymptotically) complete we mean that it is so the largest problem class that the solver naturally targets, although it may be complete on specific sub-classes. For example, {\sc SnOpt} naturally targets continuous NLPs and it is incomplete on nonconvex NLPs, and therefore on QCQ, but becomes complete for CCC. In general, all complete methods for a problem $P$ must be complete for any problem $Q \subseteq P$, while a complete method for $P$ might be incomplete for $R \supset P$.

@@

Local search methods typically require as an input a solution $x'$ and attempt to improve it---either towards feasibility, or towards optimality, or both---using only information that is available in a neighborhood of $x'$. In general, local search methods are incomplete. As remarked above, however, local search methods deployed on a convex problem behave like complete methods, possibly via devices that allow one to find a feasible starting solution if there is one (like what was used to be called ``phase -1'' in the simplex method).

Most local search methods for *C* are iterative in nature, and need a starting point $x^0$ as input. The ($k+1$)-st iterate is obtained as $x^{k+1} = x^k + \alpha_k d^k$, where $\alpha_k$ (a scalar) is the {\it step length}, and $d^k$ (a vector) is the {\it search direction}, e.g.~belonging to a tangent manifold and having a negative directional derivative at $x^k$, in order to improving optimality while reducing infeasibility. Alternatively, $x^{k+1} = x^k + d^k(\alpha_k)$ where $d^k(\cdot)$ is an arc satisfying $d^k(0) = 0$. Local search methods for *I* are also iterative, but since defining a useful notion of descent direction is harder in presence of integer variables, $d^k$ and $\alpha_k$ are usually computed in different ways, depending on the application at hand.

The solvers in Table \ref{t:solvers} which implement incomplete methods for continuous NLPs (a problem class containing QCQ) are {\sc ConOpt}, {\sc IpOpt}, {\sc Minos}, {\sc SnOpt} and {\sc KNitro}\todo{check capitalization of solver names}. The methods are local search algorithms and are essentially of the following types: %Note that traditionally {\sc KNitro} used to be a local (nonconvex) NLP solver.
%All of these solvers implement essentially three types of local search methods for continuous NLPs:
\todo{sequential linear programming, sequential quadratic programming, trust-region}
\begin{itemize}
 \item active set methods ({\sc ConOpt}, {\sc Minos}, {\sc SnOpt}) \cite{Dost09};
 %
 \item interior point methods ({\sc IpOpt}, {\sc KNitro}) \cite{Wright97};
 %
 \item projected gradient methods \cite{CalaMore87:mp,ChenGui13:coap}.
\end{itemize} \todo{doesn't Knitro additionally implement an active set or SQP algorithm?}
%
Active set and interior point methods have been defined for *C* but also for general (continuous) NLPs. In fact, all of the solvers named above target this larger problem class.
